---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "##",
    dpi = 250,
    fig.path = "man/images/"
)
```

# iebabynames: Full baby name data for Ireland

```{r, include=FALSE}
library(iebabynames)
```


## Description

Full baby name data (1964--2023) for Ireland, gathered from the [Central Statistics Office](https://www.cso.ie/en/interactivezone/visualisationtools/babynamesofireland/).

The package contains the dataset `iebabynames` with `r  format(nrow(iebabynames), big.mark = ",")` observations on six variables: `year`, `sex`, `name`, `n`, `rank`, and `prop`, and `prop_sex`. Due to confidentiality reasons, only names with 3 or more instances in the relevant year are included.

The package can be used to explore patterns of baby names in Ireland over time. The dataset `iebabynames` is also very suitable for filtering, summarising and plotting variables in workshops or lectures.

The structure of the package follows the [**babynames**](https://cran.r-project.org/web/packages/babynames/index.html) package by Hadley Wickham and the [**ukbabynames**](https://mine-cetinkaya-rundel.github.io/ukbabynames/) package by Mine Çetinkaya-Rundel, Thomas Leeper, and Nicholas Goguen-Compagnoni. 


## Installation

The package is hosted on GitHub and not available at CRAN. To install the latest development version:

```{r eval = FALSE}
if (!require("remotes")) {
    install.packages("remotes")
}
remotes::install_github("stefan-mueller/iebabynames") 
```

## Demonstration

```{r, message=FALSE}
# load packages
library(iebabynames)
library(dplyr)
library(ggplot2)
library(scales)

# set ggplot2 theme
theme_set(theme_bw())

head(iebabynames)
```

### Get most popular names in 2023

```{r}
dat_top_2023 <- iebabynames %>% 
    filter(year == "2023") %>% 
    group_by(sex) %>% 
    top_n(n = 10, wt = -rank) # get top 10

dat_top_2023
```



### Plotting the 10 most frequent male and female names across the entire period

```{r,message=FALSE, fig.width=6, fig.height=6}
iebabynames_top <- iebabynames %>% 
    group_by(sex, name) %>% 
    summarise(n_total = sum(n)) %>% 
    top_n(n = 10, wt = n_total)

ggplot(iebabynames_top, aes(x = n_total,
                            y = reorder(name, n_total), 
                            fill = sex)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = name), nudge_x = -2000, 
              hjust = "right",
              colour = "white") +
    scale_x_continuous(labels = scales::comma_format()) +
    scale_fill_manual(values = c("#83D0F5", "#71B294")) +
    labs(title = "Frequency of Baby Names in Ireland (1964–2023)", 
         x = "Number of Babies",
         y = NULL) +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.title = element_blank(),
          legend.position = "bottom") 
```


### Inspecting the development of selected names

```{r,message=FALSE, fig.width=6.5, fig.height=3}
ggplot(data = filter(iebabynames, name %in% c("John", "Mary")),
       aes(x = year, y = prop)) +
    geom_smooth(se = FALSE) +
    scale_x_continuous(breaks = c(seq(1963, 2023, 10))) +
    scale_y_continuous(labels = scales::percent,
                       breaks = c(seq(0, 0.06, 0.01))) +
    geom_point(alpha = 0.4) +
    facet_wrap(~name) +
    labs(x = NULL, y = "Percentage of Babies") 
```


### Explore different variants of names

```{r,message=FALSE, fig.width=6, fig.height=3}
iebabynames_variants <- iebabynames %>% 
    filter(name %in% c("Aoife", "Aoibhe", "Eva",
                       "Eve"))

ggplot(data = iebabynames_variants,
       aes(x = year, y = n)) +
    geom_smooth(se = FALSE) +
    scale_x_continuous(breaks = c(seq(1963, 2023, 20))) +
    geom_point(alpha = 0.4) +
    facet_wrap(~name, nrow = 1) +
    labs(x = NULL, y = "Frequency") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

The [website of the Central Statistics Office](https://www.cso.ie/en/interactivezone/visualisationtools/babynamesofireland/) includes an interactive interface that allows you to plot the frequency and rank of custom names.


## How to cite

Stefan Müller (2024). _iebabynames: Ireland Baby Names, 1964--2023_. R package version 0.2.4. URL: http://github.com/stefan-mueller/iebabynames.

If you use the data, please also cite the CSO website: https://www.cso.ie/en/interactivezone/visualisationtools/babynamesofireland/.

## Issues

Please file an issue (with a bug, wish list, etc.) [via GitHub](https://github.com/stefan-mueller/iebabynames/issues).
